### 2019-07-09 pubs/paper/sections/02-introduction.tex
Many revisions and fixes.

\documentclass[../main.tex]{subfiles}
\begin{document}

@@

An automatic theorem prover (henceforth \ATP) is a program that
intends to prove conjectures from axioms and inference rules of some
logical system. In the last decades, \ATPs are fast becoming a key
instrument in different disciplines and real applications (\eg,
verifying a railway interlocking system, an operating-system kernel,
or a pseudo-random number generator). Since some
programming errors have been found in these programs (see, for
example, \cite{Keller2013,Bohme2011,Fleury2014}), researchers and
users from academy and industry have shown an increased interest to
\emph{formally} prove the validity of \ATPs' results.

In order to give confidence to the \ATP users many of these systems
have started to include in their outputs the full derivations
associated to the proved theorems. However, existing research
recognizes that in many cases these derivations encode non-trivial
reasoning hard to reconstruct and therefore hard to
verify~\cite{paulson2007source,Keller2013}.

Proof-reconstruction addresses this problem. Since many \ATPs are
poor-documented, this problem becomes in mostly cases a reverse
engineering task to verify the prover reasoning. The usual is the
reconstruction is made by another and not by the developers of the
\ATP. Therefore, the presentation of the derivations generated by the
prover plays an important role in proof-reconstruction.

To verify such automatically generated derivation by the prover,
it is convenient to have them in a
consistent format, that is, a full script describing the derivation
step-by-step with  exhaustive details and without ambiguities. For
example, for classical propositional logic (henceforth \CPL) from a
list of at least forty \ATPs---available from the Web service
\name{SystemOnTPTP} of the TPTP World\footnote{\url{http://www.cs.miami.edu/~tptp/}.}---just few of
them show their proofs.

One approach to address the proof-reconstruction problem is
proving each deduction of the prover, the \emph{source} system,
with a formalization of the prover reasoning in a proof-assistant, the
\emph{target} system. The target system is the
proof \emph{checker} in charge to verify the source system reasoning
for each derivation.
These proof-assistants allow us to formalize the logical system used
in the proofs, \ie, logical constants, axioms, inference rules,
hypotheses, and theorems.
A proof-reconstruction tool provides such an
integration, translating the derivation generated by the prover into
the formalism of the proof-assistant.

@@

We describe a formal reconstruction of proofs generated by the \Metis
prover~\cite{hurd2003first}---our source system---in Martin-Löf type
theory~\cite{Nordstrom-Petersson-Smith-1990}. We
formalize the subset of the \Metis inference rules for the
propositional logic fragment using a syntactical treatment.
The \Metis reasoning was formalized in
\Agda~\cite{agdateam}---our target system---in two
libraries~\cite{AgdaProp,AgdaMetis} and we implemented a
proof-reconstruction tool named \Athena~\cite{Athena} written in
\Haskell that is able to generate \Agda proof-terms for \Metis
derivations. During writing this document, our formalization helps to report
two programming errors in \Metis.
A bug\footnote{Issue No.\,2 at \url{https://github.com/gilith/metis/issues/2}}
in the printing of the derivation and a soundness
bug\footnote{Issue No.\,4 at \url{https://github.com/gilith/metis/issues/4}} in the stripping of the goal.

This paper has been organized in the following way.  In
Section~\ref{sec:type-theory}, some limitations of type theory are
discussed from our proof-reconstruction point of view.  In
Section~\ref{sec:metis-language-and-proofs}, we introduce the \Metis
prover.  In Section~\ref{sec:proof-reconstruction}, we show our
approach to reconstruct \Metis derivations.  Related work is described in
Section~\ref{sec:related-work}.  Conclusions and suggestions for
future work are presented in Section~\ref{sec:conclusions}.

@@

The proof-reconstruction tool \Athena was tested with \prg{GHC}~v8.6.5.
Both libraries, \verb!agda-prop! and \verb!agda-metis! were tested with
\Agda~v2.6.0.1 and \Agda standard library~v1.1.
\Athena jointly with \verb!agda-prop! and \verb!agda-metis! are able
to reconstruct propositional proofs of \Metis~v2.4.20180810.
We successfully reconstruct around eighty theorems in classical propositional
logic from the \TPTP collection~\cite{Prieto-Cubides2017} to test
the developments and the formalization presented in this research.

>>>>

\documentclass[../paper.tex]{subfiles}
\begin{document}

@@

An automatic theorem prover (henceforth \ATP) is a program that
intends to prove conjectures from axioms and inference rules of some
logical system. In the last decades, \ATPs are quickly becoming a key
instrument in many disciplines and real applications (\eg, verifying a
railway interlocking system, an operating-system kernel, or a
pseudo-random number generator). Since some programming errors have
been found in these programs (see, for example,
\cite{Keller2013,Bohme2011,Fleury2014}), researchers and users from
academia and industry have shown an increased interest in
\emph{formally} proving the validity of \ATPs' results.

In order to give confidence to the users of the \ATP, many of these
systems have started to include in their outputs the full derivations
associated to the proved theorems. However, existing research
recognizes that in many cases these derivations encode non-trivial
reasoning hard to reconstruct and therefore hard to
verify~\cite{paulson2007source,Keller2013}.

Proof-reconstruction addresses this problem. Since many \ATPs are
poor-documented, this problem becomes a reverse
engineering task to verify the prover reasoning. Usually, is the
reconstruction is made by another and not by the developers of the
\ATP. Therefore, the presentation of the derivations generated by the
prover plays an important role in proof-reconstruction.

To verify such automatically generated derivations by the prover, it
is convenient to have them in a consistent format, that is, a full
script describing the derivation step-by-step with  exhaustive details
and without ambiguities. For example, from a list of at least forty
\ATPs for classical propositional logic (henceforth \CPL) ---available
from the Web service \name{SystemOnTPTP} of the TPTP
World\footnote{\url{http://www.cs.miami.edu/~tptp/}.}--- just few of
them show proofs when they find the conjecture of the problem is a 
theorem.

One approach to address the proof-reconstruction problem is
proving each deduction of the prover, the \emph{source} system,
with a formalization of the prover reasoning in a proof-assistant, the
\emph{target} system. The target system is the
\emph{proof-checker} in charge of verifying the source system reasoning
for each derivation.
These proof-assistants allow us to formalize the logical system used
in the proofs, \ie, logical constants, axioms, inference rules,
hypotheses, and theorems.
A proof-reconstruction tool provides such an
integration, translating the derivation generated by the prover into
the formalism of the proof-assistant.

@@

We describe a formal reconstruction of proofs generated by the \Metis
prover~\cite{hurd2003first}---our source system---in Martin-Löf type
theory~\cite{Nordstrom-Petersson-Smith-1990}. To do so, we only  take
into account the syntax rather than the semantics of the subset of the
\Metis inference rules for the propositional logic fragment. The
\Metis reasoning was formalized in \Agda~\cite{agdateam}---our target
system---in two libraries~\cite{AgdaProp,AgdaMetis} and we implemented
a proof-reconstruction tool named \Athena~\cite{Athena} written in
\Haskell that is able to generate \Agda proof-terms for \Metis
derivations.  In the course of this work, we uncovered and reported
two programming errors in the ATP \Metis: a bug\footnote{Issue No.\,2
at \url{https://github.com/gilith/metis/issues/2}} in the printing of
the derivation and a soundness bug\footnote{Issue No.\,4 at
\url{https://github.com/gilith/metis/issues/4}} in the stripping of
the goal.

This paper is organized in the following way: in
Section~\ref{sec:type-theory}, some limitations of type theory are
discussed from our proof-reconstruction point of view.  In
Section~\ref{sec:metis-language-and-proofs}, we introduce the \Metis
prover.  In Section~\ref{sec:proof-reconstruction}, we show our
approach to reconstruct \Metis derivations.  Related work is described
in Section~\ref{sec:related-work}.  Conclusions and suggestions for
future work are presented in Section~\ref{sec:conclusions}.

@@

The proof-reconstruction tool \Athena was tested with
\prg{GHC}~v8.6.5. Both libraries, \verb!agda-prop! and
\verb!agda-metis! were tested with \Agda~v2.6.0.1 and \Agda standard
library~v1.1. \Athena jointly with \verb!agda-prop! and
\verb!agda-metis! are able to reconstruct propositional proofs of
\Metis~v2.4.20180810. We successfully reconstructed eighty theorems in
classical propositional logic from the \TPTP
collection~\cite{Prieto-Cubides2017} to test the developments and the
formalization presented in this research.

### 2019-07-09 pubs/paper/sections/03-type-theory.tex
Many revisions and fixes.

Type theory is a formalism for the foundation of mathematics, it has
became in a key instrument to study logic and proof theory that
follows the same basis of constructive mathematics where the witness
of a statement is \emph{everything}. In that direction, we could say
the main actors in type theory are the statements and their proofs.

By following the Curry-Howard correspondence (see, for
example,~\cite{Wadler-2015}), a formula corresponds to a \emph{type},
and one proof of that formula is a \emph{term} of the correspondent
type. Therefore, inhabitants types are such formulas with proofs, they are theorems.

Since type theory is a formal system, we have a syntax and a set of
derivation rules. These rules enables us to derive a kind of
conclusions called \emph{judgments}. A judgment is another way to say
that a term has a certain type. Each term has associated a derivation,
we refer to this derivation as proof-terms.

@@

  We write the type judgments as $a : \type{A}$ to denote that the
  term~$a$ is of type~$\type{A}$.

@@

We produce derivation trees inside the judgments using the derivation
rules. Since type theory can be seen as typed $λ$-calculus with
dependent function types, evaluation of $λ$-terms also called
\emph{normalization} of the proof-terms, is a process of reductions
with the system inference rules.

Therefore, the proof verification task becomes in type theory as
verifying that the proof-term has the correspondent type of the
theorem. We know this process as \emph{type-checking}.
This feature of type theory allow us to verify a proof generated by
an \ATP by reconstructing its proof-term to type-check the proof.
For such a purpose, we use a proof-assistant based on type theory like
\Agda to delegate this task. % However, this approach has some weak points
discussed in~\cite{Bezem2002}. Some limitations from the type theory
point of view for proof-reconstruction is described in
Section~\ref{ssec:structural-recursion} and
Section~\ref{ssec:intuitionistic-logic}.

@@

To reconstruct \Metis inference rules in type theory, we observed
that some rules or their inner functions are implemented by
\emph{general recursive} functions (see a definition of these functions in~\cite{Bove2002}).

Functions defined by a general recursion can not be directly
translated in type theory since it is not a guarantee they terminate.
For that reason, we follow the technique described
in~\cite{Bertot2004} to avoid termination problems by modifying the
recursive functions to be \emph{structurally recursive}.

A recursive function is structurally recursive if it calls itself
with only structurally smaller arguments~\cite{Abel2002}.
General recursive functions can be rewrote into structurally
recursive functions by  using for instance, the
\emph{bounded recursion} technique. For a yet another methods, we refer the
reader to \cite{Coquand1992,Abel2002,Bove2005}.

The bounded technique defines a new structural recursive function
based on the general recursive function by adding an argument. The
new argument is the \emph{bound}, a natural number given by the
function complexity.
In other words, the added argument will represent the number of times
the function needs to call itself to get the expected outcome.

\begin{notation}
  We use \Prop for the type of propositions.
  A proposition is an expression of indivisible propositional variables (\eg, symbols $φ₀, φ₁, \dots$), and the logic constants:
  $⊥$, $⊤$, the binary connectives ($∧, ∨, ⇒)$, and the negation
  $(\neg)$. We use the inductive definition for propositions presented in~\cite{VanDalen1994}.

  We write the syntactical equality between two propositions φ and ψ using the symbol (≡) as $φ ≡ ψ$. For this equality, we assume the reflexivity, symmetry and
  transitivity properties.

  The type of natural numbers is called \Nat,
  and it is defined as usual, \ie, $\zero$ and $\suc$ are its data
  constructors. We use names and symbols for the arithmetic
  operations as usual (\eg, $+$, $-$, $*$).
  We use syntax sugar for $\zero$, $\suc$,
  $\suc~\zero$, $\dots$, with the decimal representation~$0$, $1$,
  $2$, $\dots$ as well.
\end{notation}

Hence, one approach to define a structural recursive function based
on a general
recursive function $f₀ : \type{A} \to \type{B}$ is to
formulate a new
function $f₁ : \type{A} \to \Nat \to \type{B}$ where all
of its recursive calls are done on smaller arguments.

@@

\begin{equation}
\label{eq:uh-definition}
\begin{aligned}
&\hspace{.495mm}\fuh_{0} : \Prop → \Prop\\
&\begin{array}{lll}
  \fuh_{0} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{0}~((φ₁ ∧ φ₂) ⇒ φ₃)\\
  \fuh_{0} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fuh_{0}~(φ₁ ⇒ φ₂) ∧ \fuh_{0}~(φ₁ ⇒ φ₃)\\
  \fuh_{0} &φ                &= φ.
\end{array}
\end{aligned}
\end{equation}

In~\eqref{eq:uh-definition}, the first two equations in the $\fuh_{0}$
function definition are no structurally recursive.
Note that the formula used in the recursive calls are not subformulas
of the formula in the function argument.

% TODO
% , the first two equations
% call on arguments no structurally smaller.
% For instance, in the first equation,
% $$\fuh_{0} (φ₁ ⇒ (φ₂ ⇒ φ₃)) = \fuh_{0}~((φ₁ ∧ φ₂) ⇒ φ₃),$$
% the function $\fuh_{0}$ calls on the argument $(φ₁ ∧ φ₂) ⇒ φ₃$ argument.
% However, this formula is not a subformula of $φ₁ ⇒ (φ₂ ⇒ φ₃)$.

Therefore, one way to define $\fuh_{0}$ in type
theory is to define a new function using a bounded recursion.

\begin{equation}
\label{eq:uh-structured}
\begin{aligned}
&\hspace{.495mm} \fuh_{1} : \Prop → \Nat → \Prop\\
&\begin{array}{llll}
\fuh_{1} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &(\suc~n) &= \fuh_{1}~((φ₁ ∧ φ₂) ⇒ φ₃)~n\\
\fuh_{1} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &(\suc~n) &= \fuh_{1}~(φ₁ ⇒ φ₂)~n ∧ \fuh_{1}~(φ₁ ⇒ φ₃)~n\\
\fuh_{1} &φ &n &= φ.
\end{array}
\end{aligned}
\end{equation}

We bounded the recursion calls of the $\fuh_{0}$ function using its
\emph{complexity measure}.

The complexity measure of a function is the number of steps the
function takes to finish. Since the $\fuh_{0}$ function is recursive,
we can define its complexity measure by defining a recursive function
instead of a closed formula for such a number. However, this function
must to be structurally recursive as well and can be defined by
following the pattern-matching cases of its definition.

Therefore, we define in~\eqref{eq:uh-complexity} the complexity
measure function of $\fuh_{0}$ by assigning the number of steps to
finish for each pattern-matching case.

@@

We have used the complexity measure definition for a formula defined in~\cite{Agudelo-Agudelo2017} to define other complexity measures
in this paper.

\begin{equation}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{equation}

Following the technique mentioned above, we define the
function~$\fuh$, the structural
recursive definition of the function~$\fuh_{0}$.
\begin{align*}
  \label{eq:uh}
  \begin{split}
  &\fuh : \Prop → \Prop\\
  &\fuh~φ~ = \fuh_{1}~φ~(\fuh_{cm}~φ).
  \end{split}
\end{align*}

@@

% For a more complete treatment of general recursion, and how to define
% well-founded and well-structural recursive functions, we refer the
% reader to~\cite{Bove2005}.

@@

Type theory and intuitionistic logic are accompanying theories since
they are based on the same philosophical basis of constructive
mathematics. Proving propositions in these theories demands a witness
construction for the proof. Nonetheless, classical logic does not
always have constructive proofs since some of the proofs are stated
by refutation.

@@

We formalize the syntactical consequence relation of \CPL by an
inductive family \verb!_⊢_! with two indexes, a set of propositions (the
premises) and a proposition (the conclusion), that is, $Γ ⊢ φ$
represents that there is derivation with conclusion $φ : \Prop$ from
the set of premises $Γ : \List~\Prop$. We implemented
in~\cite{AgdaProp} the syntactical consequence relation in a similar
way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and
elimination rules for connectives and the PEM axiom as the valid
inference rules in Fig.~\ref{fig:CPL-inference-rules}.

>>>>

Type theory is a formalism for the foundation of mathematics. It has
become a key instrument to study logic and proof theory, and follows
the same basis of constructive mathematics where the witnesses of a
statement, the proofs, have a fundamental role as same as their
statement.

Following the Curry-Howard correspondence (see, for
example,~\cite{Wadler-2015}), a formula corresponds to a \emph{type},
and one proof of that formula is a \emph{term} of the correspondent
type. Therefore, inhabited types are such formulas with proofs, they
are theorems.

Type theory is a formal system with a syntax and a set of derivation
rules. These rules enables us to derive conclusions called
\emph{judgments} which can be seen as derivation
trees~\cite{Bezem2002}.  Each term has associated a derivation and we
refer to these derivations as \emph{proof-terms}.

@@

  We write the typing judgments as $a : \type{A}$ to denote that the
  term~$a$ is of type~$\type{A}$.

@@

Type theory can be seen as typed $λ$-calculus with
dependent function types, evaluation of $λ$-terms also called
\emph{normalization} of the proof-terms, is a process of reductions
with the system inference rules. Therefore, the proof verification
task becomes in type theory as verifying that the proof-term has the
correspondent type of the theorem. We refer to this process as
\emph{type-checking}.

This feature of type theory allow us to verify a proof generated by an
\ATP by reconstructing its proof-term to type-check the proof. For
such a purpose, we use a proof-assistant based on type theory, \Agda,
to delegate this task as in~\cite{Bezem2002}. Some limitations
of the type theory point-of-view for proof-reconstruction is described
in Section~\ref{ssec:structural-recursion} and
Section~\ref{ssec:intuitionistic-logic}.

@@

To reconstruct \Metis inference rules in type theory, we observed that
some rules or their inner functions are implemented by \emph{general
recursive} functions (see the definition of these functions
in~\cite{Bove2002}).

Functions defined by a general recursion can not be directly
translated in type theory since it is not guaranteed that they terminate.
For that reason, we follow the technique described
in~\cite{Bertot2004} to avoid termination problems by modifying the
recursive functions to be \emph{structurally recursive}.

A recursive function is structurally recursive if it calls itself with
only structurally smaller arguments~\cite{Abel2002}. General recursive
functions can sometimes be rewritten into structurally recursive
functions by  using for instance, the \emph{bounded recursion}
technique or other methods as in \cite{Coquand1992,Abel2002,Bove2005}.

The bounded recursion technique defines a new structural recursive function
based on the general recursive function by adding an argument. The
new argument is the \emph{bound}, a natural number given by the
function complexity.
In other words, the added argument will represent the number of times
the function needs to call itself to get the expected outcome.

We define \Prop for the type of propositions. A proposition is an
expression of indivisible propositional variables (\eg, symbols $φ₀,
φ₁, \dots$), and the logic constants: $⊥$, $⊤$, the binary connectives
($∧, ∨, ⇒)$, and the negation $(\neg)$. We use the inductive
definition for propositions presented in~\cite{VanDalen1994}. The
syntactical equality between two propositions φ and ψ is written using
the symbol (≡) as $φ ≡ ψ$. This equality between propositions is
precisely the Identity type $\mathsf{Id}_{\Prop}(φ, ψ)$ in Martin-Löf
type theory. Therefore, it holds the reflexivity, symmetry and
transitivity property for equality. The type of natural numbers is
called \Nat, and it is defined as usual, \ie, $\zero$ and $\suc$ are
its data constructors. We use names and symbols for the arithmetic
operations as usual (\eg, $+$, $-$, $*$). We use syntax sugar for
$\zero$, $\suc$, $\suc~\zero$, $\dots$, with the decimal
representation~$0$, $1$, $2$, $\dots$ as well.

Hence, one approach to define a structural recursive function based on
a general recursive function $f₀ : \type{A} \to \type{B}$ is to
formulate a new function $f₁ : \type{A} \to \Nat \to \type{B}$ where
all of its recursive calls are done with \emph{smaller} values on its
second argument~\cite{Abel2002}.

@@

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:uh-definition}
  \begin{aligned}
  &\hspace{.495mm}\fuh_{0} : \Prop → \Prop\\
  &\begin{array}{lll}
  \fuh_{0} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{0}~((φ₁ ∧ φ₂) ⇒ φ₃)\\
  \fuh_{0} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fuh_{0}~(φ₁ ⇒ φ₂) ∧ \fuh_{0}~(φ₁ ⇒ φ₃)\\
  \fuh_{0} &φ                &= φ.
\end{array}
  \end{aligned}
  \end{empheq}

In~\eqref{eq:uh-definition}, the first two equations in the $\fuh_{0}$
function definition are not structurally recursive. As a consequence,
we could not use this function in \Agda which would report a
non-terminating error. This happens because the formula used in the
recursive calls are not subformulas of the input formula in the
function argument. Therefore, one way to define $\fuh_{0}$ in type
theory is to define a new function using a bounded recursion.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation*}
\label{eq:uh-structured}
    \begin{aligned}
    &\hspace{.495mm} \fuh_{1} : \Prop → \Nat → \Prop\\
    &\begin{array}{llll}
    \fuh_{1} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &(\suc~n) &= \fuh_{1}~((φ₁ ∧ φ₂) ⇒ φ₃)~n\\
    \fuh_{1} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &(\suc~n) &= \fuh_{1}~(φ₁ ⇒ φ₂)~n ∧ \fuh_{1}~(φ₁ ⇒ φ₃)~n\\
    \fuh_{1} &φ &n &= φ.
    \end{array}
  \end{aligned}
\end{empheq}

We bounded the recursion calls of the $\fuh_{0}$ function using its
\emph{complexity measure}. We refer to the complexity measure of a
function as the number of steps the function takes to finish. Since
the $\fuh_{0}$ function is recursive, we can define its complexity
measure by defining a recursive function instead of a closed formula
for such a number. However, this function must to be structurally
recursive as well, and it can be defined by following the
pattern-matching cases of its definition. Thus,  the complexity
measure of the $\fuh_{0}$ function can be defined as follows.

@@

% We have used the complexity measure definition for a formula defined
% in~\cite{Agudelo-Agudelo2017} to define other complexity measures in
% this paper.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation*}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{empheq}

Finally, we are able to define the $\fuh$ function which is the
structural recursive definition of the function~$\fuh_{0}$ also
suitable to use in \Agda.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation*}
 \begin{aligned}
  \label{eq:uh}
  \begin{split}
  &\hspace{.495mm}\fuh_{cm} : \Prop → \Prop\\
  &\fuh~φ~ = \fuh_{1}~φ~(\fuh_{cm}~φ).
  \end{split}
\end{aligned}
\end{empheq}

@@

@@

Type theory is foundation for constructive mathematics, and its
internal language is the intuitionistic logic. Proving propositions in
this theory demands a witness construction for the proof.
Traditionally, for example in classical logic, a witness is not always
needed as we can prove arguments by contraction, named \emph{reductio
ad absurdum}.

@@

We formalize the syntactical consequence relation of \CPL by an
inductive family \verb!_⊢_! with two indexes, a set of propositions
(the premises) and a proposition (the conclusion), that is, $Γ ⊢ φ$
represents that there is derivation with conclusion $φ : \Prop$ from
the set of premises $Γ : \List~\Prop$. We implemented
in~\cite{AgdaProp} the syntactical consequence relation in a similar
way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included in Fig.~\ref{fig:CPL-inference-rules} the valid
inference rules: structural rules like \emph{weaken}, formation and
elimination rules of the logical connectives and to reason
classically, we add the \abbre{PEM} axiom.

### 2019-07-09 pubs/paper/sections/04-metis-language-and-proofs.tex
Many revisions and fixes.

%
\documentclass[../main.tex]{subfiles}
\begin{document}

@@

The \TSTP language is an output language for derivations used by
\Metis and other ATPs~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}.
A \TSTP  derivation generated by \Metis encodes a natural
deduction proof using only as valid deduction steps the inference rules
in Fig.~\ref{tab:agda-metis-table}.
These derivations are directed acyclic graphs, trees of refutations.
Each node stands for an application of an inference rule to the formulas
in the parent nodes. The leaves in the tree represent formulas in the given
\TPTP input. Each node is labeled with a name of its inference rule (\eg, \canonicalize).
Each edge in the tree links a premise with one conclusion.

>>>>

%
\documentclass[../paper.tex]{subfiles}
\begin{document}

@@

The \TSTP language is an output language for derivations used by
\Metis and other ATPs~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}.
A \TSTP  derivation generated by \Metis encodes a natural
deduction proof using as valid deduction steps the inference rules
in Fig.~\ref{tab:agda-metis-table}.
These derivations are directed acyclic graphs, trees of refutations.
Each node stands for an application of an inference rule to the formulas
in the parent nodes. The leaves in the tree represent formulas in the given
\TPTP input. Each node is labeled with a name of its inference rule (\eg, \canonicalize).
Each edge in the tree links a premise with one conclusion.
